### 2019-04-10 paper/paper.tex
added loc for paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2018

% ready for submission
% \usepackage{neurips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
     \usepackage[final]{neurips_2018}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{MINST Kaggle Digit Recognizer: Contrasting the Randomized Forest andMLP Neural Network}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
   Andrew Osborne \\
   \texttt{amo004@uark.edu} \\
   \And
   Josh Price \\
   \texttt{jdp024@uark.edu} \\
   \AND
   April Walker \\
   \texttt{adw027@uark.edu} \\
   \\
   University of Arkansas, \\
   Fayetteville, AR, 72701, USA
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
  In order to tackle the classic MINST dataset, our team developed a randomized forest and a multi-layer perception (MLP) classifier using the \verb+sklearn+ Python library. While our RF was able to achieve about 6\% test error with only 10 trees, increasing the number of trees soon developed diminishing returns. Cross validation never breached 97\% accuracy on test data, suggesting an RF is not well suited to solve such a high dimensional problem. In order to optimize our MLP NN, we explored the use of stochastic and Adam gradient descent, however found the later consistently outperformed the former. In order to balance accuracy vs. complexity, our cross validation method suggests the use of 1 hidden layer and 256 nodes. Our final Kaggle submission resulted in a correct prediction rate of 97.90\%, suggesting our NN is well-trained.    
\end{abstract}

\section{The MINST Dataset}

I'll probably talk about the considerations going into this project and the reason behind our chosen models. It might also be nice to just give a general background of the dataset that will set this up?

\section{Randomized Forest}

Here I will give a basic whatever on RF's that Dr. Zhang will already know. "It's like a decision tree with extra steps that gives you 'good enough' results." -Someone

\subsection{Implementation}

This is probably not the best title for this section

\subsection{Cross-Validation and Results}

\subsection{Future Considerations}
Because an RF is probably so so, but there are probably similar models that work better on problems w/ high levels of dimensionality.

\section{Multi-Layer Perceptron Classifier}
This will be much more interesting to talk about

\subsection{Contrasting Gradient Descent Algorithms}
Because it will probably be good to talk about why Adam's so good, basically because it used the components of a lot of other good methods.

\subsection{Cross-Validation and Results}

\section{Conclusions}
MLP NN's are noice, Los Bobo's is better.

\clearpage

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references. {\bf Remember that you can use more than eight
  pages as long as the additional pages contain \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2019-04-12 src/IO.py
feeble attempt at editing for readability.. I really need a style guide

@@ -5,6 +5,9 @@ import matplotlib.pyplot as plt
 
 # skl expect n_samples x n_features array and n_samples array
 def readData(filename,test=None):
+    """
+        This simply reads the data csv files and handles test, train cases
+    """
     if test == True:
         data=np.genfromtxt(filename,dtype=float,delimiter=',',skip_header=1)
         #np.random.shuffle(data)

### 2019-04-12 src/proj_test.py

@@ -2,6 +2,7 @@ import numpy as np
 import joblib as jl
 import matplotlib as mpl
 mpl.use("Agg")
+# generate plots over an ssh connection
 import matplotlib.pyplot as plt
 from sklearn.neural_network import MLPClassifier
 from sklearn.model_selection import cross_val_score, GridSearchCV
@@ -13,33 +14,50 @@ import sys
 import pandas as pds
 
 def readModel(filename):
+    """
+        reads a model file from a python pickle
+        expect a neural net, not a random forest
+    """
     clf = joblib.load(filename)
     return clf
 
 def genCVplots(filename):
+    """
+        generates crossvalidation plots 
+        from the model learned in proj_train.py
+        will plot mean cross validation score against 
+        alpha (regularization) term.
+    """
     clf = readModel(filename)
     dat = pds.DataFrame(clf.cv_results_)
+    
+    # print out the pandas dataFrame from the learned model
     print(dat)
+
     #fig = dat.plot('param_alpha','mean_test_score',kind='line',subplots=True,logx=True)[0].get_figure()
     #fig.savefig("dat.png")
     #params = {'a':[0.5,0.1,0.001,0.0001],'hl':[(128,128),(512,),(256,)],'s':['sgd','adam']}
     
+    # stupid solution to a stupid problem
     hl128x128_solS = []
     hl128x128_solA = []
     hl256_solS = []
     hl256_solA = []
     hl512_solS = []
     hl512_solA = []
+    # perm contains [alpha, mean CV score, mean train score, mean fit time]
+    # in each component. 
     perm = [hl128x128_solS,hl128x128_solA,hl256_solS,hl256_solA,hl512_solS,hl512_solA]
     for x in range(0,6):
         for y in range(0,4):
             i = x + 6*y
             perm[x].append([dat['param_alpha'][i],dat['mean_test_score'][i],dat['mean_train_score'][i],dat['mean_fit_time'][i]])
     
-    #print(np.array(perm)[0,:,:])
     pl = np.array(perm)
 
 
+    # these arrays are used for the clumsy labeling of rows and 
+    # columns in the final plot
     r = ['2 HL, 128 nodes','1 HL, 256 nodes','1 HL, 512 nodes']
     c = ['sgd','adam']
     for x in range(0,6):
@@ -64,6 +82,14 @@ def genCVplots(filename):
     #pds.plotting.parallel_coordinates(dat,'Names')
 
 def prediction(modelfile, datafile, outfile):
+    """
[…]

### 2019-04-12 src/proj_train.py

@@ -6,14 +6,28 @@ from sklearn.preprocessing import minmax_scale
 from IO import readData
 import os
 
-
+"""
+    This file will train a neural net on the test data,
+    on two cores. This takes a lot of time to train
+"""
 
 
 if __name__=="__main__":
+    # parameter grid for GridSearchCV
     parameters = {'alpha':[0.5,0.1,0.001,0.0001],'hidden_layer_sizes':[(128,128),(256,),(512,)], 'solver':['sgd','adam']}
+    
+    # the parameters set here are held constant in the subsequent grid search
     nn = MLPClassifier(verbose=10,activation='relu',learning_rate='adaptive',tol=1e-4)
+
+    # this trains several models on the above defined grid, and saves some data
     clf = GridSearchCV(nn,parameters,verbose = 10, n_jobs = 2,cv=4)
+
+    # read data, and scale it
     X_pre,y = readData('../data/train.csv')
     X = minmax_scale(X_pre)
+
+    # train the model
     clf.fit(X,y)
+
+    # save the model to a pickle
     joblib.dump(clf,"clf.joblib")

### 2019-04-12 src/train_dt.py

@@ -9,11 +9,22 @@ import os
 
 
 if __name__=="__main__":
+    # parameter grid for the grid search
     parameters = {'n_estimators':[10,50,100,300],'min_samples_split':[2,4,8,16]}
+
+    # parameters set here are held constant in grid search
     dt = RandomForestClassifier(verbose=10,max_features='sqrt')
+
+    # trian model on earlier defined 
     clf = GridSearchCV(dt,parameters,verbose = 10, n_jobs = 2,cv=4)
+
+    # read and scale features and labels
     X_pre,y = readData('../data/train.csv')
     X = minmax_scale(X_pre)
+
+    # train the model, takes much less time to train than the neural net
     clf.fit(X,y)
+
+    # write file to python pickle: kind of a big file
     joblib.dump(clf,"dt.joblib")
 

### 2019-04-13 img/scripts/render.py
inverted img render

@@ -16,8 +16,9 @@ def render(filename,i):
     img = np.array(A[i,:],copy=True)
     print(img.shape)
     img = img.reshape(28,28)
+    img = 255 - img
     print(img.shape)
-    plt.imshow(img,cmap='gray')
+    plt.imshow(img, cmap="gray", vmin=0, vmax=255)
     plt.savefig("img" + str(i)+"render"+ ".png")
 
 

### 2019-04-14 paper/paper.tex
paper updates

@@
 \usepackage{microtype}      % microtypography
+\usepackage{graphicx}
 
-\title{MINST Kaggle Digit Recognizer: Contrasting the Randomized Forest andMLP Neural Network}
+\title{MINST Kaggle Digit Recognizer: Contrasting the Random Forest and MLP Neural Network}
 
@@
 \begin{abstract}
-  In order to tackle the classic MINST dataset, our team developed a randomized forest and a multi-layer perception (MLP) classifier using the \verb+sklearn+ Python library.
-While our RF was able to achieve about 6\% test error with only 10 trees, increasing the number of trees soon developed diminishing returns.
-Cross validation never breached 97\% accuracy on test data, suggesting an RF is not well suited to solve such a high dimensional problem.
+  In order to tackle the classic MINST dataset, our team implemented a random forest (RF) and a multi-layer perception (MLP) classifier using the \verb+sklearn+ Python library.
+As expected from a high-dimensional problem such as digit recognition, our RF preformed sub-optimally.
+Contrasting the accuracy between our training data and testing data suggest severe overfitting.
+Our final Kaggle submission reached a correct prediction rate of 96.66\%.
 In order to optimize our MLP NN, we explored the use of stochastic and Adam gradient descent, however found the later consistently outperformed the former.
@@
 Our final Kaggle submission resulted in a correct prediction rate of 97.90\%, suggesting our NN is well-trained.
+While our MLP preformed considerably better than our RF, the training took approximately 60 times as long, implying the RF may be a desirable option when speed is of utmost importance.
 \end{abstract}
@@
 \section{The MINST Dataset}
+The MNIST digit database is a classic starting point for individuals wanting to learn the basics of computer vision and get hands on experience with machine learning algorithms.
+The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine.
+The pre-flattened Kaggle dataset provides you with 42,000 training examples and 28,000 testing examples.
+Each 28x28 pixel image is represented by a 784 length vector with each element containing some integer between 0 and 255 representing the lightness or darkness of that pixel.
+Additionally, the training dataset provides the correct labels for each image.
+These images can be reshaped and rendered from the provided dataset as shown in Figure 1.
+\begin{figure}[h]
+  \centering
+    \begin{minipage}[t]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img786render.png}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img98render.png}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{img6render.png}
+    \end{minipage}
+  \caption{Image Renderings from the MNIST Dataset}
+\end{figure}
+\section{Random Forest}
+A Random Forest (RF) is a common starting algorithm for those new to machine learning.
+This is in part due to it's easy of use, but also due to it's flexibility.
+While in this assignment it is used for classification, it can be also be used in regression problems.
+Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
+Rather than searching for the very best split feature, it chooses the best of a random subset.
+Unlike the standard RF algorithm, \verb+sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, rather than picking a classification based on a majority vote.
 
-I'll probably talk about the considerations going into this project and the reason behind our chosen models.
-It might also be nice to just give a general background of the dataset that will set this up?
-
-\section{Randomized Forest}
-
-Here I will give a basic whatever on RF's that Dr. Zhang will already know. "It's like a decision tree with extra steps that gives you 'good enough' results." -Someone
+These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.
+Generally speaking, overfitting can be negated by simple adding more trees to your model.
+This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical.
 
 \subsection{Implementation}
-
-This is probably not the best title for this section
+The 784 length vector of input data is normalized to range from 0 to 1 using \verb+sklearn+'s \verb+preprocessing.minmax_scale+.
+In order to improve our model's predictive power, we contrasted the results from our hyperparameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).
+Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leafs.
+The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.
+Once the code was prepared, training took approximately 45 minutes.
+\begin{figure}[h]
+    \centering
+    \begin{minipage}[t]{0.45\textwidth}
+        \centering
+        \textbf{(A)}
+        \includegraphics[width=\textwidth]{CVscoreVsLeafSize.png}
+    \end{minipage}
+    \begin{minipage}[t]{0.45\textwidth}
+        \centering
+        \textbf{(B)}
+        \includegraphics[width=\textwidth]{CVtestVsLeafSize.png}
+    \end{minipage}
+\caption{(A,B) Cross Validation of Hyperparameters for (A) training data and (B) testing data by contrasting mean cross validation score and $\log_2(minleafsize)$ giving the minimum leaf threshold as a measure of complexity.
+The scores for varying numbers of trees are shown.}
+\end{figure}
+\section{Random Forest}
 
 \subsection{Cross-Validation and Results}
-
-\subsection{Future Considerations}
-Because an RF is probably so so, but there are probably similar models that work better on problems w/ high levels of dimensionality.
+Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
+In order to visualize the results, Figure 2 contrasts the prediction rates of the training and test datasets.
+The $\log_2(minleafsize)$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data.
+Adding trees has a similar positive affect on both datasets, however, also decreases exponentially as more trees are added.
+These results suggest that our model suffers from extreme levels of overfitting.
+As expected, adding trees does improve our models predictive power but isn't the most practical for our dataset.
+Our Kaggle submission gave a prediction rate of 96.6\%
 
@@
 
+\subsection{Implementation}
+
 \subsection{Cross-Validation and Results}
@@
 \section{Conclusions}
-MLP NN's are noice, Los Bobo's is better.
 

### 2019-04-14 paper/paper.tex
1st rough draft

@@
 In order to balance accuracy vs.
-complexity, our cross validation method suggests the use of 1 hidden layer and 256 nodes.
+complexity, our cross validation method suggests the use of 1 hidden layer and 512 nodes.
 Our final Kaggle submission resulted in a correct prediction rate of 97.90\%, suggesting our NN is well-trained.
@@
 The MNIST digit database is a classic starting point for individuals wanting to learn the basics of computer vision and get hands on experience with machine learning algorithms.
-The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine.
+The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine [6].
 The pre-flattened Kaggle dataset provides you with 42,000 training examples and 28,000 testing examples.
@@
 \end{figure}
+\section{GridSearchCV}
+For both machine learning algorithms, training and cross-validation was done using \verb+sklearn+'s \verb+model_selection.GridSearchCV+.
+The grid search takes in a grid of the hyper-parameters the user wishes to contrast and exhaustively considers all possible parameter combinations.
+Rather than train the data on the entire training dataset, \verb+GridSearchCV+ partitions the data into $k$ sections then trains the data on $k-1$ of the sections, leaving the last partition as a pseudo-test dataset.
+The mean cross-validation score (CVS) is the averaged prediction rate over all $k$ segments of the training data, and the hyper-parameter combination with the best mean CVS is chosen [4].
+
 \section{Random Forest}
@@
 This is in part due to it's easy of use, but also due to it's flexibility.
-While in this assignment it is used for classification, it can be also be used in regression problems.
+While in this assignment it is used for classification, it can be also be used in regression problems [2].
 Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
-Rather than searching for the very best split feature, it chooses the best of a random subset.
-Unlike the standard RF algorithm, \verb+sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, rather than picking a classification based on a majority vote.
+Rather than searching for the very best split feature, it chooses the best of a random subset.\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
 
 These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.
-Generally speaking, overfitting can be negated by simple adding more trees to your model.
-This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical.
+Generally speaking, overfitting can be negated by simple adding more trees to your model [2].
+This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical [5].
 
@@
 The 784 length vector of input data is normalized to range from 0 to 1 using \verb+sklearn+'s \verb+preprocessing.minmax_scale+.
-In order to improve our model's predictive power, we contrasted the results from our hyperparameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).
+In order to improve our model's predictive power, we contrasted the results from our hyper-parameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).
 Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leafs.
 The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.
-Once the code was prepared, training took approximately 45 minutes.
+Once the code was prepared, training took approximately 45 seconds.
 \begin{figure}[h]
@@
     \end{minipage}
-\caption{(A,B) Cross Validation of Hyperparameters for (A) training data and (B) testing data by contrasting mean cross validation score and $\log_2(minleafsize)$ giving the minimum leaf threshold as a measure of complexity.
+\caption{(A,B) Cross Validation of Hyper-parameters for (A) training data and (B) testing data by contrasting mean cross validation score and $\log_2(minleafsize)$ giving the minimum leaf threshold as a measure of complexity.
 The scores for varying numbers of trees are shown.}
 \end{figure}
-\section{Random Forest}
 
@@
 Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
-In order to visualize the results, Figure 2 contrasts the prediction rates of the training and test datasets.
+In order to visualize the results, Figure 2 contrasts the mean CVS and prediction rates of the training and test datasets.
 The $\log_2(minleafsize)$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data.
@@
 As expected, adding trees does improve our models predictive power but isn't the most practical for our dataset.
-Our Kaggle submission gave a prediction rate of 96.6\%
+Our Kaggle submission of 300 trees and a minimum leaf number of 2 gave a prediction rate of 96.6\%
 
 \section{Multi-Layer Perceptron Classifier}
-This will be much more interesting to talk about
+A Multi-layer Perceptron (MLP) is a type of deep neural network (NN) that is trained using backpropogation.
+An MLP consists of at least one hidden layer (HL), an input (features) layer, and an output (targets) layer [2].
 
 \subsection{Contrasting Gradient Descent Algorithms}
-Because it will probably be good to talk about why Adam's so good, basically because it used the components of a lot of other good methods.
+A key component of implementing any model using backpropogation is choosing an appropriate gradient descent (GD) algorithm.
+For our MLP, we decided to contrast stochastic gradient descent (SGD) and Adaptive Moment Estimation (Adam) gradient descent developed by Kingma and Lei Ba [3].
+As we've discussed in class, stochastic gradient descent approximates the gradient by considering a single training example at a time.
+Adam gradient descent is actually an optimized SGD which computes adaptive learning rates for each parameter.
+This method stores exponentially decaying averages of the past gradient $\nabla_\theta f_t(\theta_{t-1})$ and square of the gradient $\nabla_\theta f_t(\theta_{t-1})^2$ as estimates of the first and second moment of the gradient (mean $m_t$ and variance $v_t$ respectively).
+The exponential decay rates $\beta_1$ and $\beta_2$ are initialized (generally near 1)\footnote{The developers recommended default is $\beta_1 = 0.9$ and $\beta_2 = 0.999$.
+[3]} such that the moments can be more specifically calculated using the following:
+$$m_t \leftarrow \beta_1 \cdot m_{t-1} + (1-\beta_1)\cdot \nabla_\theta f_t(\theta_{t-1}) $$
+$$v_t \leftarrow \beta_2 \cdot v_{t-1} + (1-\beta_2)\cdot \nabla_\theta f_t(\theta_{t-1})^2 $$
+With the initial $m_0$ and $v_0$ set to 0. In order to correct for bias, the following bias-corrected moments are then computed:
+$$\hat{m_t} \leftarrow m_t/(1-\beta_1^t)$$
+$$\hat{v_t} \leftarrow v_t/(1-\beta_2^t)$$
+With $\beta_1^t$ and $\beta_2^t$ indicating $\beta_1$ and $\beta_2$ to the power of $t$.
+Given some $\alpha$ and $\epsilon$ as regularization parameters, the parameters $\theta_t$ are then updated using the following:
+$$\theta_t \leftarrow \theta_{t-1} - \alpha \cdot \hat{m_t}/(\sqrt{\hat{v_t}+\epsilon})$$
+This process is repeated until $\theta$ converges [3].
 
 \subsection{Implementation}
+In addition to contrasting GD algorithms, the number of hidden nodes were varied between 128, 256, and 512; the number of hidden layers were either 1 or 2; the regularization term $\alpha$ was varied at 0.5, 0.1, 0.001, and 0.0001.
+The other regularization parameter's for Adam GD were kept at the developer's (Kingma and Lei Ba) recommended defaults: $\epsilon = 10^{-8}$, $\beta_1=0.9$, and $\beta_2=0.999$ [3].
+While the learning rate was not varied, an adaptive rate was chosen which divides the current learning rate by 5 with a starting value of $0.001$ [4].
+One the code was prepared, training took approximately 46 minutes.
 
 \subsection{Cross-Validation and Results}
+Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
+In order to visualize the results, Figure 2 contrasts the mean CVS and prediction rates of the training and test datasets.
+A measure of complexity is given by $-log_{10}$ meaning lower $\alpha$ values correspond to higher complexity.
+As expected, Adam gradient descent was the clear winner across the board.
+Note that the smaller $\alpa$ value (that is the larger $-\log_{10}(\alpha)$ value) almost consistently improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
+Examples most obvious example of this is at 1 HL and 256 nodes in (B).
+Our "winning" model with 1 HL and 512 nodes preformed rather consistently between training and testing datasets.
+The Kaggle submission of this model gave a prediction rate of 97.90\%.
+
+\begin{figure}[h]
+    \centering
+    \begin{minipage}[t]{0.45\textwidth}
+        \centering
+        \textbf{(A)}
+        \includegraphics[width=\textwidth]{TrainscoreVsAlpha.png}
+    \end{minipage}
+    \begin{minipage}[t]{0.45\textwidth}
+        \centering
+        \textbf{(B)}
+        \includegraphics[width=\textwidth]{CVscoreVsAlpha.png}
+    \end{minipage}
+\caption{(A,B) Cross Validation of Hyper-parameters for (A) training data and (B) testing data by contrasting the mean CVS in (A) and the prediction accuracy in (B) with $-\log_{10}(\alpha)$ giving a measure of complexity.
+For both (A) and (B), the left colomn gives the results for SGD, and the right colmn gives the results for Adam GD.}
+\end{figure}
 
 \section{Conclusions}
+As expected, our MLP was better suited for handwritten digit recognition.
+It is possible that data processing to reduce dimensional would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
+One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding addition trees would begin to bridge this gap and further reduce it's run time speed.
 
@@
 \section*{References}
-
-References follow the acknowledgments. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references. {\bf Remember that you can use more than eight
-  pages as long as the additional pages contain \emph{only} cited references.}
 \medskip
@@
 
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
+[1] Bernard, S., Adam, S., & Heutte, L.
+(2007).
+\textit{Using Random Forests for Handwritten Digit Recognition}.
+Ninth International Conference on Document Analysis and Recognition (ICDAR 2007) Vol 2.
+doi:10.1109/icdar.2007.4377074
 
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
+[2] Hastie, T., Tibshirani, R., & Friedman, J. H. (2017). The Elements of Statistical Learning: Data Mining, Inference, and Prediction. New York, NY: Springer.
 
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
+[3] Kingma, D. P., & Lei Ba, J. (2015). 
+\textit{Adam: A Method for Stochastic Optimization}. Conference Paper at ICLR. Retrieved from https://arxiv.org/abs/1412.6980.
+
+[4] Pedregosa \textit{et al} (2011) \textit{Scikit-learn: Machine Learning in Python}. JMLR 12, pp. 2825-2830.
+
+[5] Robnik-Šikonja, M. (2004). \textit{Improving Random Forests.} Machine Learning: ECML 2004 Proceedings, Springer, Berlin, 359-370.
+
+[6] Y.
+LeCun \textit{et al} (1995) \textit{Learning Algorithms For Classification: A Comparison On Handwritten Digit Recognition}, in Oh, J.
+H.
+and Kwon, C.
+and Cho, S.
+(Eds), Neural Networks: The Statistical Mechanics Perspective, 261-276, World Scientific.
 

### 2019-04-15 paper/paper.tex
minor edits to abstract and other things

@@
 Our final Kaggle submission resulted in a correct prediction rate of 97.90\%, suggesting our NN is well-trained.
-While our MLP preformed considerably better than our RF, the training took approximately 60 times as long, implying the RF may be a desirable option when speed is of utmost importance.
+While our MLP preformed considerably better than our RF, the training took approximately 60 times as long, implying the RF may be a desirable option when speed in training is of utmost importance.
+On the other hand, RF classification time is much greater than that of a NN, which is a nontrivial consideration.
 \end{abstract}
@@
 Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
-Rather than searching for the very best split feature, it chooses the best of a random subset.\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
+Our random forest is a ``highly random'' random forest which splits data with no discretion to promote model volatility.\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
 
@@
 In order to improve our model's predictive power, we contrasted the results from our hyper-parameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).
-Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leafs.
+Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leaves.
 The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.
@@
 In order to visualize the results, Figure 2 contrasts the mean CVS and prediction rates of the training and test datasets.
-The $\log_2(minleafsize)$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data.
+The $\log_2(minleafsize)$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data, a symptom of overfitting.
 Adding trees has a similar positive affect on both datasets, however, also decreases exponentially as more trees are added.
@@
 While the learning rate was not varied, an adaptive rate was chosen which divides the current learning rate by 5 with a starting value of $0.001$ [4].
-One the code was prepared, training took approximately 46 minutes.
+One the code was prepared, training took approximately 46 minutes per model.
+Total training time took approximately 20 hours.
 
@@
 As expected, our MLP was better suited for handwritten digit recognition.
-It is possible that data processing to reduce dimensional would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
-One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding addition trees would begin to bridge this gap and further reduce it's run time speed.
+It is possible that data processing to reduce dimensionality would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
+One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding addition trees would begin to bridge this gap and further improve it's classification accuracy.
 

### 2019-04-21 Presentation/Presentation.tex
Paper changes
Add files via upload

%% Based on a TeXnicCenter-Template by Gyorgy SZEIDL.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------
%
\documentclass[titlepage,leqno]{beamer}%
\usetheme{Antibes}%
\usecolortheme{rose}%
\setbeamerfont{footnote}{size=\tiny}
%
%----------------------------------------------------------
% This is a sample document for the LaTeX Slides Class
% Class options
%       --  Body text point size (normalsize) is 27 (default)
%           and can not be adjusted to any other value.
%       --  Paper size:  letterpaper (8.5x11 inch, default)
%                        a4paper, a5paper, b5paper,
%                        legalpaper, executivepaper
%       --  Orientation (portrait is the default):
%                        landscape
%       --  Quality:     final(default), draft
%       --  Title page:  titlepage, notitlepage
%       --  Columns:     onecolumn (default), [twocolumn is not avalible]
%       --  Equation numbering (equation numbers on the right is the default)
%                        leqno (equation numbers on the left)
%       --  Displayed equations (centered is the default)
%                    fleqn (flush left)
%
%  \documentclass[a4paper,fleqn]{slides}
%
%  The slides are separated from each other by the slide
%  environment, see below:
%
%%%
% Some helpful information:
%
% You can typeset \emph{Emphasized text}.
%
% You can also typeset  \textbf{Bold}, \textit{Italics},
% \textsl{Slanted} and \texttt{Typewriter} text. Roman fonts are not
% available.
%
% Point size can be changed by making use of the {\tiny tiny},
% {\scriptsize scriptsize}, {\footnotesize footnotesize}, {\small
% small}, {\normalsize normalsize}, {\large large}, {\Large Large},
% {\LARGE LARGE}, {\huge huge} and {\Huge Huge} commands.
%
% The numbered equation
% \begin{equation}
% u_{tt}-\Delta u+u^{5}+u\left|  u\right|  ^{p-2}=0\text{ in }\mathbf{R}%
% ^{3}\times\left[  0,\infty\right[  .\label{eqn1}%
% \end{equation}
% is automatically numbered as equation \ref{eqn1}.
%%%

\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}%
\usepackage{adjustbox}%
\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{verbatim}

%%  \usepackage{bibentry} 
%%  \usepackage{natbib}
%------------------------------------------------------------------
\hfuzz5pt % Don't bother to report overfull boxes < 5pt
\pagestyle{plain}
%%% --------------------------------------------------------------
\begin{document}
\title{MINST Kaggle Digit Recognizer: Contrasting the Random Forest and MLP Neural Network}
\author{%
   Andrew Osborne \hspace{2.75cm} Josh Price\\
   \texttt{amo004@uark.edu} \hspace{1.75cm} \texttt{jdp024@uark.edu} \\
   \vspace{4mm}
   April Walker \\
   \texttt{adw027@uark.edu}
}
\institute{%
   University of Arkansas, \\
   Fayetteville, AR, 72701, USA
	}

\date{4/23/2019}
\maketitle
%-----------------------------------------------------------------
\begin{frame}{Presentation Outline}
	\begin{itemize}
		\item The MINST Dataset
		\item GridSearchCV
		\item Random Forest
		\begin{itemize}
			\item Implementation
			\item Results
		\end{itemize}
		\item Multi-Layer Perception Classifier
		\begin{itemize}
			\item Contrasting Gradient Descent Algorithms
			\item Implementation
			\item Results
		\end{itemize}
		\item{Conclusions}
	\end{itemize}
\end{frame}
% ----------------------------------------------------------------
\begin{frame}{The MINST Dataset}
%The MNIST digit database is a classic starting point for individuals wanting to learn the basics of computer vision and get hands on experience with machine learning algorithms. The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine [6]. The pre-flattened Kaggle dataset provides you with 42,000 training examples and 28,000 testing examples. Each 28x28 pixel image is represented by a 784 length vector with each element containing some integer between 0 and 255 representing the lightness or darkness of that pixel. Additionally, the training dataset provides the correct labels for each image. These images can be reshaped and rendered from the provided dataset as shown in Figure 1. 
\begin{figure}[h]
  \centering
    \begin{minipage}[t]{0.3\textwidth}
        \includegraphics[width=\textwidth]{img786render.png}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \includegraphics[width=\textwidth]{img98render.png}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \includegraphics[width=\textwidth]{img6render.png}
    \end{minipage}
  \caption{Image Renderings from the MNIST Dataset}
\end{figure}
\end{frame}
%-----------------------------------------------------------------
\begin{frame}{The MINST Dataset}
\begin{itemize}
	\item 70,000 $28\times 28$ pixel grey-scale images of handwritten numbers, 0 through 9.
	\item Each image is represented by a vector of length 784, with each element taking a value between 0 and 255 to represent lightness/darkness of the pixel
	\item Pre-flattened Kaggle dataset
	\begin{itemize}
		\item 42,000 training examples
		\item 28,000 testing examples
	\end{itemize}
\end{itemize}
%The MNIST digit database is a classic starting point for individuals wanting to learn the basics of computer vision and get hands on experience with machine learning algorithms. The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine [6]. The pre-flattened Kaggle dataset provides you with 42,000 training examples and 28,000 testing examples. Each 28x28 pixel image is represented by a 784 length vector with each element containing some integer between 0 and 255 representing the lightness or darkness of that pixel. Additionally, the training dataset provides the correct labels for each image. These images can be reshaped and rendered from the provided dataset as shown in Figure 1. 
\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{GridSearchCV}
\begingroup
\small
For both machine learning algorithms, training and cross-validation was done using \verb+sklearn+'s \verb+model_selection.GridSearchCV+.

\vspace{3.75mm}

The grid search takes in a grid of the hyper-parameters the user wishes to contrast and exhaustively considers all possible parameter combinations.

\vspace{3.75mm}

\verb+GridSearchCV+ partitions the data into $k$ sections then trains the data on $k-1$ of the sections, leaving the last partition as a pseudo-test dataset.

\vspace{3.75mm}

The mean cross-validation score (CVS) is the averaged prediction rate over all $k$ segments of the training data, and the hyper-parameter combination with the best mean CVS is chosen [4]. 
\endgroup
\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{Random Forest}

Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.

\vspace{3.75mm}

Our random forest is a ``highly random'' random forest which splits data with no discretion to promote model volatility.

\vspace{3.75mm}

\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Random Forest}

These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.

\vspace{3.75mm}

Generally, overfitting can be negated by adding more trees to your model [2].

\vspace{3.75mm}

This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical [5].

\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{Random Forest}
\small
\textbf{Implementation}

Data normalized to values in [0,1] using \verb+sklearn+'s \verb+preprocessing.minmax_scale+.

\vspace{3.75mm}

In order to improve our model's predictive power, we contrasted the results from our hyper-parameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).

\vspace{3.75mm}

Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leaves. The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.

\vspace{3.75mm}

Once the code was prepared, training took approximately 45 seconds.

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Random Forest}

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \textbf{(A)}
        \includegraphics[width=\textwidth]{CVscoreVsLeafSize.png}
    \end{minipage}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \textbf{(B)}
        \includegraphics[width=\textwidth]{CVtestVsLeafSize.png}
    \end{minipage}
\caption{\footnotesize
(A,B) Cross Validation of Hyper-parameters for (A) training data and (B) testing data by contrasting mean cross validation score and $\log_2(minleafsize)$ giving the minimum leaf threshold as a measure of complexity. The scores for varying numbers of trees are shown.}
\end{figure}

\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{Random Forest}
\textbf{Cross-Validation and Results}

The $\log_2(\text{minleafsize})$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data, a symptom of overfitting.

\vspace{3.75mm}

Adding trees has a similar positive affect on both datasets, however, also decreases exponentially as more trees are added. These results suggest that our model suffers from extreme levels of overfitting.

\vspace{3.75mm}

Our Kaggle submission of 300 trees and a minimum leaf number of 2 gave a prediction rate of 96.6\%

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Multi-Layer Perceptron Classifier}
\textbf{Contrasting Gradient Descent Algorithms}

For our MLP, we contrast stochastic gradient descent (SGD) and Adaptive Moment Estimation gradient descent (Adam GD), developed by Kingma and Lei Ba [3].

\vspace{3.75mm}

\begin{itemize}
	\item SGD approximates the gradient by considering a single training example at a time.
	\item Adam GD is an optimized SGD which computes adaptive learning rates for each parameter.
\end{itemize}

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Multi-Layer Perceptron Classifier}
\tiny
\textbf{Adam Parameter Estimation}

\begin{itemize}
	\item Exponentially decaying averages of the past gradient $\nabla_\theta f_t(\theta_{t-1})$ and 
	square of the gradient $\nabla_\theta f_t(\theta_{t-1})^2$ are stored as estimates of the first 
	and second moment of the gradient (mean $m_t$ and variance $v_t$ respectively).
	\item The exponential decay rates $\beta_1$ and $\beta_2$ are initialized (generally near 1)\footnote{The 
	developers recommended default is $\beta_1 = 0.9$ and $\beta_2 = 0.999$. [3]} such that the moments can be 
	more specifically calculated using the following:
	\begin{align*}
	m_t \leftarrow& \beta_1 \cdot m_{t-1} + (1-\beta_1)\cdot \nabla_\theta f_t(\theta_{t-1})\\
	v_t \leftarrow& \beta_2 \cdot v_{t-1} + (1-\beta_2)\cdot \nabla_\theta f_t(\theta_{t-1})^2 
	\end{align*}
	\item With the initial $m_0$ and $v_0$ set to 0. In order to correct for bias, the following bias-corrected 
	moments are then computed:
	\begin{align*}
	\hat{m_t} \leftarrow& m_t/(1-\beta_1^t)\\
	\hat{v_t} \leftarrow& v_t/(1-\beta_2^t)
	\end{align*}
	\item With $\beta_1^t$ and $\beta_2^t$ indicating $\beta_1$ and $\beta_2$ to the power of $t$. 
	Given some $\alpha$ and $\epsilon$ as regularization parameters, the parameters $\theta_t$ are then 
	updated using the following:
	\begin{align*}
	\theta_t \leftarrow& \theta_{t-1} - \alpha \cdot \hat{m_t}/(\sqrt{\hat{v_t}+\epsilon})
	\end{align*}
	\item This process is repeated until $\theta$ converges [3].
\end{itemize}

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Multi-Layer Perceptron Classifier}
\scriptsize
\textbf{Implementation}

In addition to contrasting GD algorithms we also varied:
\begin{itemize}
	\item the number of hidden nodes between 128, 256, and 512
	\item the number of hidden layers either 1 or 2
	\item $\alpha$ at 0.5, 0.1, 0.001, and 0.0001
\end{itemize}
\medskip
Adam GD regularization parameters were initialized at the developer's (Kingma and Lei Ba) recommended defaults:
\begin{itemize}
	\item $\epsilon = 10^{-8}$ 
	\item $\beta_1=0.9$ 
	\item $\beta_2=0.999$ [3]
\end{itemize}
\medskip
The learning rate was not varied, instead an adaptive rate was chosen which divides the current learning rate by 5 with a starting value of $0.001$ [4].

\medskip

Once the code was prepared, training took approximately 46 minutes per model. Total training time took approximately 20 hours.

\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{Multi-Layer Perceptron Classifier}
\scriptsize
\textbf{Cross-Validation and Results}
\begin{itemize}
	\item Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
	\item The measure of complexity is given by $-\log_{10}(\alpha)$, meaning lower $\alpha$ values 
	correspond to higher complexity.
\end{itemize}
\medskip
As expected, Adam gradient descent was the clear winner across the board.

Note:
\begin{itemize}
	\item Smaller $\alpha$ values (that is the larger $-\log_{10}(\alpha)$ values) almost consistently 
	improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
	\item The most obvious example of this is at 1 HL and 256 nodes in (B). 
\end{itemize} 
\medskip

Our "winning" model with 1 HL and 512 nodes preformed rather consistently between training and testing datasets.

\medskip

The Kaggle submission of this model gave a prediction rate of 97.90\%.

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Multi-Layer Perceptron Classifier}

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \textbf{(A)}
        \includegraphics[width=\textwidth]{TrainscoreVsAlpha.png}
    \end{minipage}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \textbf{(B)}
        \includegraphics[width=\textwidth]{CVscoreVsAlpha.png}
    \end{minipage}
\caption{\scriptsize
(A,B) Cross Validation of Hyper-parameters for (A) training data and (B) testing data by contrasting the mean CVS in (A) and the prediction accuracy in (B) with $-\log_{10}(\alpha)$ giving a measure of complexity. For both (A) and (B), the left colomn gives the results for SGD, and the right colmn gives the results for Adam GD.}
\end{figure}

\end{frame}
%-----------------------------------------------------------------
\begin{frame}{Conclusions}

As expected, our MLP was better suited for handwritten digit recognition.

\vspace{3.75mm}

It is possible that data processing to reduce dimensionality would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.

\vspace{3.75mm}

One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding additional trees would begin to bridge this gap and further improve it's classification accuracy.

\end{frame}
%-----------------------------------------------------------------
\begin{frame}[fragile]{References}

\tiny

[1] Bernard, S., Adam, S., \& Heutte, L. (2007). \textit{Using Random Forests for Handwritten Digit Recognition}. Ninth International Conference on Document Analysis and Recognition (ICDAR 2007) Vol 2. doi:10.1109/icdar.2007.4377074

\vspace{3mm}

[2] Hastie, T., Tibshirani, R., \& Friedman, J. H. (2017). The Elements of Statistical Learning: Data Mining, Inference, and Prediction. New York, NY: Springer.

\vspace{3mm}

[3] Kingma, D. P., \& Lei Ba, J. (2015). 
\textit{Adam: A Method for Stochastic Optimization}. Conference Paper at ICLR. Retrieved from https://arxiv.org/abs/1412.6980.

\vspace{3mm}

[4] Pedregosa \textit{et al} (2011) \textit{Scikit-learn: Machine Learning in Python}. JMLR 12, pp. 2825-2830.

\vspace{3mm}

[5] Robnik-Šikonja, M. (2004). \textit{Improving Random Forests.} Machine Learning: ECML 2004 Proceedings, Springer, Berlin, 359-370.

\vspace{3mm}

[6] Y. LeCun \textit{et al} (1995) \textit{Learning Algorithms For Classification: A Comparison On Handwritten Digit Recognition}, in Oh, J. H. and Kwon, C. and Cho, S. (Eds), Neural Networks: The Statistical Mechanics Perspective, 261-276, World Scientific.

\end{frame}
%-----------------------------------------------------------------
\begin{frame}
	\begin{center}
	\begingroup
	\huge
	Thank You!\\
	\endgroup
	\end{center}
\end{frame}
%-----------------------------------------------------------------
\end{document}

### 2019-04-21 paper/paper.tex
Paper changes
Add files via upload

@@
 A Random Forest (RF) is a common starting algorithm for those new to machine learning.
-This is in part due to it's easy of use, but also due to it's flexibility.
+This is in part due to it's ease of use, but also due to it's flexibility.
 While in this assignment it is used for classification, it can be also be used in regression problems [2].
@@
 These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.
-Generally speaking, overfitting can be negated by simple adding more trees to your model [2].
+Generally speaking, overfitting can be negated by simply adding more trees to your model [2].
 This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical [5].
@@
 In order to visualize the results, Figure 2 contrasts the mean CVS and prediction rates of the training and test datasets.
-The $\log_2(minleafsize)$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data, a symptom of overfitting.
+The $\log_2(\text{minleafsize})$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data, a symptom of overfitting.
 Adding trees has a similar positive affect on both datasets, however, also decreases exponentially as more trees are added.
@@
 In addition to contrasting GD algorithms, the number of hidden nodes were varied between 128, 256, and 512; the number of hidden layers were either 1 or 2; the regularization term $\alpha$ was varied at 0.5, 0.1, 0.001, and 0.0001.
-The other regularization parameter's for Adam GD were kept at the developer's (Kingma and Lei Ba) recommended defaults: $\epsilon = 10^{-8}$, $\beta_1=0.9$, and $\beta_2=0.999$ [3].
+The other regularization parameters for Adam GD were kept at the developer's (Kingma and Lei Ba) recommended defaults: $\epsilon = 10^{-8}$, $\beta_1=0.9$, and $\beta_2=0.999$ [3].
 While the learning rate was not varied, an adaptive rate was chosen which divides the current learning rate by 5 with a starting value of $0.001$ [4].
-One the code was prepared, training took approximately 46 minutes per model.
+Once the code was prepared, training took approximately 46 minutes per model.
 Total training time took approximately 20 hours.
@@
 In order to visualize the results, Figure 2 contrasts the mean CVS and prediction rates of the training and test datasets.
-A measure of complexity is given by $-log_{10}$ meaning lower $\alpha$ values correspond to higher complexity.
+A measure of complexity is given by $-\log_{10}(\alpha)$ meaning lower $\alpha$ values correspond to higher complexity.
 As expected, Adam gradient descent was the clear winner across the board.
-Note that the smaller $\alpa$ value (that is the larger $-\log_{10}(\alpha)$ value) almost consistently improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
-Examples most obvious example of this is at 1 HL and 256 nodes in (B).
+Note that the smaller $\alpha$ values (that is the larger $-\log_{10}(\alpha)$ values) almost consistently improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
+The most obvious example of this is at 1 HL and 256 nodes in (B).
 Our "winning" model with 1 HL and 512 nodes preformed rather consistently between training and testing datasets.
@@
 It is possible that data processing to reduce dimensionality would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
-One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding addition trees would begin to bridge this gap and further improve it's classification accuracy.
+One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding additional trees would begin to bridge this gap and further improve it's classification accuracy.
 

### 2019-04-23 Presentation/Presentation.tex
Paper/Presentation changes

@@
 \setbeamerfont{footnote}{size=\tiny}
+\setbeamercolor{footnotet}{bg=white}
 %
@@
 \usepackage{amsmath}%
+\usepackage{nopageno}%
+\usepackage{framed}%
+\usepackage{animate}%
 \usepackage{amsfonts}%
@@
 \begin{itemize}
-	\item 70,000 $28\times 28$ pixel grey-scale images of handwritten numbers, 0 through 9.
+	\item 70,000 $28\times 28$ pixel grey-scale images of handwritten numbers, 0 through 9
+	\vspace{3.5mm}
 	\item Each image is represented by a vector of length 784, with each element taking a value between 0 and 255 to represent lightness/darkness of the pixel
+	\vspace{3.5mm}
 	\item Pre-flattened Kaggle dataset
@@
 \small
-For both machine learning algorithms, training and cross-validation was done using \verb+sklearn+'s \verb+model_selection.GridSearchCV+.
+For both machine learning algorithms, training and cross-validation was done using \verb+sklearn+'s \verb+model_selection.GridSearchCV+
 
@@
 
-The grid search takes in a grid of the hyper-parameters the user wishes to contrast and exhaustively considers all possible parameter combinations.
+\verb+GridSearchCV+ partitions the data into $k$ sections then trains the data on $k-1$ of the sections, leaving the last partition as a pseudo-test dataset
 
@@
 
-\verb+GridSearchCV+ partitions the data into $k$ sections then trains the data on $k-1$ of the sections, leaving the last partition as a pseudo-test dataset.
+The mean cross-validation score (CVS) is the averaged prediction rate over all $k$ segments of the training data, and the hyper-parameter combination with the best mean CVS is chosen
 
-\vspace{3.75mm}
+\vspace{3.7mm}
 
-The mean cross-validation score (CVS) is the averaged prediction rate over all $k$ segments of the training data, and the hyper-parameter combination with the best mean CVS is chosen [4]. 
 \endgroup
@@
 
-Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
+Our random forest is a ``highly random'' random forest which splits data with no discretion to promote model volatility
 
@@
 
-Our random forest is a ``highly random'' random forest which splits data with no discretion to promote model volatility.
+\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional majority vote approach
 
-\vspace{3.75mm}
+\vspace{3.7mm}
 
-\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
+These models are generally very fast to train, but depending on the complexity, can suffer from slow classification time
 
-\end{frame}
-%-----------------------------------------------------------------
-\begin{frame}{Random Forest}
+\vspace{3.7mm}
 
-These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.
-
-\vspace{3.75mm}
-
-Generally, overfitting can be negated by adding more trees to your model [2].
-
-\vspace{3.75mm}
-
-This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical [5].
+High-dimensional data often poses a problem for Random Forests.
 
@@
 
-Data normalized to values in [0,1] using \verb+sklearn+'s \verb+preprocessing.minmax_scale+.
+\medskip
 
+In order to optimize our model's predictive power, we contrasted the results from the following hyper-parameters:
+\begin{itemize}
+    \item \verb+n_estimators+ (number of trees)
+    \item \verb+min_samples_split+ (minimum leaves to split node) % required to split a node
+\end{itemize}
 \vspace{3.75mm}
 
-In order to improve our model's predictive power, we contrasted the results from our hyper-parameters \verb+n_estimators+ (number of trees) and  \verb+min_samples_split+ (minimum number of leafs required to split a node).
+Once the code was prepared, training took approximately 45 seconds per model.
 
+\end{frame}
+%-----------------------------------------------------------------
+\begin{frame}[fragile]{Random Forest}
+\small
+\textbf{Implementation}
+
+\medskip
+
+In order to optimize our model's predictive power, we contrasted the results from the following hyper-parameters:
+\begin{itemize}
+    \item \verb+n_estimators+ \{10, 20, 100, 300\}
+    \item \verb+min_samples_split+ \{2,4,8,16\}
+\end{itemize}
 \vspace{3.75mm}
 
-Specifically we allowed 10, 50, 100, and 300 trees to be developed and 2, 4, 8, and 16 as the minimum number of leaves.
-The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.
-
-\vspace{3.75mm}
-
-Once the code was prepared, training took approximately 45 seconds.
+Once the code was prepared, training took approximately 45 seconds per model.
 
@@
 \begin{frame}[fragile]{Random Forest}
+\small
 \textbf{Cross-Validation and Results}
 
-The $\log_2(\text{minleafsize})$ which inversely measures the complexity suggests the more complex models preform outstandingly against the training data but have little impact on the testing data, a symptom of overfitting.
+\medskip
+
+The $\log_2(\text{minleafsize})$ which inversely measures the complexity suggests the more complex models performs outstandingly against the training data but has little impact on the testing data, a symptom of overfitting.
 
@@
 \vspace{3.75mm}
-
+\begin{framed}
+\centering
 Our Kaggle submission of 300 trees and a minimum leaf number of 2 gave a prediction rate of 96.6\%
-
+\end{framed}
 \end{frame}
@@
 \textbf{Contrasting Gradient Descent Algorithms}
+\vspace{3.75mm}
 
-For our MLP, we contrast stochastic gradient descent (SGD) and Adaptive Moment Estimation gradient descent (Adam GD), developed by Kingma and Lei Ba [3].
+A Multi-layer Perceptron (MLP) is a type of deep neural network that is trained using backpropogation.
+An MLP consists of at least one hidden layer, an input/features layer, and an output/targets layer
+\vspace{3.75mm}
+
+For our MLP, we contrast stochastic gradient descent (SGD) and Adaptive Moment Estimation gradient descent (Adam GD)
 
@@
 
-\begin{itemize}
-	\item SGD approximates the gradient by considering a single training example at a time.
-	\item Adam GD is an optimized SGD which computes adaptive learning rates for each parameter.
-\end{itemize}
-
+\end{frame}
+%-----------------------------------------------------------------
+%\begin{frame}{Multi-Layer Perceptron Classifier}
+%\tiny
+%\textbf{Adam Parameter Estimation}
+%
+%\begin{itemize}
+	%\item Exponentially decaying averages of the past gradient $\nabla_\theta f_t(\theta_{t-1})$ and
+	%square of the gradient $\nabla_\theta f_t(\theta_{t-1})^2$ are stored as estimates of the first
+	%and second moment of the gradient (mean $m_t$ and variance $v_t$ respectively).
+	%\item The exponential decay rates $\beta_1$ and $\beta_2$ are initialized (generally near 1)\footnote{The
+	%developers recommended default is $\beta_1 = 0.9$ and $\beta_2 = 0.999$. [3]} such that the moments can be
+	%more specifically calculated using the following:
+	%\begin{align*}
+	%m_t \leftarrow& \beta_1 \cdot m_{t-1} + (1-\beta_1)\cdot \nabla_\theta f_t(\theta_{t-1})\\
+	%v_t \leftarrow& \beta_2 \cdot v_{t-1} + (1-\beta_2)\cdot \nabla_\theta f_t(\theta_{t-1})^2
+	%\end{align*}
+	%\item With the initial $m_0$ and $v_0$ set to 0. In order to correct for bias, the following bias-corrected
+	%moments are then computed:
+	%\begin{align*}
+	%\hat{m_t} \leftarrow& m_t/(1-\beta_1^t)\\
+	%\hat{v_t} \leftarrow& v_t/(1-\beta_2^t)
+	%\end{align*}
+	%\item With $\beta_1^t$ and $\beta_2^t$ indicating $\beta_1$ and $\beta_2$ to the power of $t$.
+	%Given some $\alpha$ and $\epsilon$ as regularization parameters, the parameters $\theta_t$ are then
+	%updated using the following:
+	%\begin{align*}
+	%\theta_t \leftarrow& \theta_{t-1} - \alpha \cdot \hat{m_t}/(\sqrt{\hat{v_t}+\epsilon})
+	%\end{align*}
+	%\item This process is repeated until $\theta$ converges [3].
+%\end{itemize}
+%
+%\end{frame}
+%-----------------------------------------------------------------
+\begin{frame}{Multi-Layer Perceptron Classifier}
+    \textbf{Optimizing SGD}
+    \includegraphics[width=0.9\textwidth]{patho.png}
 \end{frame}
@@
 \begin{frame}{Multi-Layer Perceptron Classifier}
-\tiny
-\textbf{Adam Parameter Estimation}
-
-\begin{itemize}
-	\item Exponentially decaying averages of the past gradient $\nabla_\theta f_t(\theta_{t-1})$ and 
-	square of the gradient $\nabla_\theta f_t(\theta_{t-1})^2$ are stored as estimates of the first 
-	and second moment of the gradient (mean $m_t$ and variance $v_t$ respectively).
-	\item The exponential decay rates $\beta_1$ and $\beta_2$ are initialized (generally near 1)\footnote{The 
-	developers recommended default is $\beta_1 = 0.9$ and $\beta_2 = 0.999$. [3]} such that the moments can be 
-	more specifically calculated using the following:
-	\begin{align*}
-	m_t \leftarrow& \beta_1 \cdot m_{t-1} + (1-\beta_1)\cdot \nabla_\theta f_t(\theta_{t-1})\\
-	v_t \leftarrow& \beta_2 \cdot v_{t-1} + (1-\beta_2)\cdot \nabla_\theta f_t(\theta_{t-1})^2 
-	\end{align*}
-	\item With the initial $m_0$ and $v_0$ set to 0. In order to correct for bias, the following bias-corrected 
-	moments are then computed:
-	\begin{align*}
-	\hat{m_t} \leftarrow& m_t/(1-\beta_1^t)\\
-	\hat{v_t} \leftarrow& v_t/(1-\beta_2^t)
-	\end{align*}
-	\item With $\beta_1^t$ and $\beta_2^t$ indicating $\beta_1$ and $\beta_2$ to the power of $t$. 
-	Given some $\alpha$ and $\epsilon$ as regularization parameters, the parameters $\theta_t$ are then 
-	updated using the following:
-	\begin{align*}
-	\theta_t \leftarrow& \theta_{t-1} - \alpha \cdot \hat{m_t}/(\sqrt{\hat{v_t}+\epsilon})
-	\end{align*}
-	\item This process is repeated until $\theta$ converges [3].
-\end{itemize}
-
-\end{frame}
-%-----------------------------------------------------------------
-\begin{frame}{Multi-Layer Perceptron Classifier}
-\scriptsize
+\small
 \textbf{Implementation}
 
-In addition to contrasting GD algorithms we also varied:
+\medskip
+
+In addition to our gradient descent algorithm, the following hyper-parameters were also varied:
 \begin{itemize}
-	\item the number of hidden nodes between 128, 256, and 512
-	\item the number of hidden layers either 1 or 2
-	\item $\alpha$ at 0.5, 0.1, 0.001, and 0.0001
+	\item Hidden Nodes: 128, 256, and 512
+	\item Hidden Layers: 1 and 2
+	\item Regularization Parameter ($\alpha$): 0.5, 0.1, 0.001, and 0.0001
 \end{itemize}
 \medskip
-Adam GD regularization parameters were initialized at the developer's (Kingma and Lei Ba) recommended defaults:
-\begin{itemize}
-	\item $\epsilon = 10^{-8}$ 
-	\item $\beta_1=0.9$ 
-	\item $\beta_2=0.999$ [3]
-\end{itemize}
-\medskip
-The learning rate was not varied, instead an adaptive rate was chosen which divides the current learning rate by 5 with a starting value of $0.001$ [4].
+%Adam GD regularization parameters were initialized at the developer's (Kingma and Lei Ba) recommended defaults:
+%\begin{itemize}
+	%\item $\epsilon = 10^{-8}$
+	%\item $\beta_1=0.9$
+	%\item $\beta_2=0.999$ [3]
+%\end{itemize}
+%\medskip
+The learning rate was not varied, instead an adaptive rate was chosen %which divides the current learning rate by 5 with a starting value of $0.001
 
@@
 Once the code was prepared, training took approximately 46 minutes per model. Total training time took approximately 20 hours.
-
-\end{frame}
-%-----------------------------------------------------------------
-\begin{frame}[fragile]{Multi-Layer Perceptron Classifier}
-\scriptsize
-\textbf{Cross-Validation and Results}
-\begin{itemize}
-	\item Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
-	\item The measure of complexity is given by $-\log_{10}(\alpha)$, meaning lower $\alpha$ values 
-	correspond to higher complexity.
-\end{itemize}
-\medskip
-As expected, Adam gradient descent was the clear winner across the board.
-
-Note:
-\begin{itemize}
-	\item Smaller $\alpha$ values (that is the larger $-\log_{10}(\alpha)$ values) almost consistently 
-	improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
-	\item The most obvious example of this is at 1 HL and 256 nodes in (B). 
-\end{itemize} 
-\medskip
-
-Our "winning" model with 1 HL and 512 nodes preformed rather consistently between training and testing datasets.
-
-\medskip
-
-The Kaggle submission of this model gave a prediction rate of 97.90\%.
 
@@
 %-----------------------------------------------------------------
+\begin{frame}[fragile]{Multi-Layer Perceptron Classifier}
+\scriptsize
+\textbf{Cross-Validation and Results}
+
+\smallskip
+
+\begin{itemize}
+	\item Training and cross-validation was done using \verb+model_selection.GridSearchCV+.
+	\item The measure of complexity is given by $-\log_{10}(\alpha)$, meaning lower $\alpha$ values
+	correspond to higher complexity.
+\end{itemize}
+\medskip
+As expected, Adam gradient descent was the clear winner across the board.
+
+Note:
+\begin{itemize}
+	\item Smaller $\alpha$ values (that is the larger $-\log_{10}(\alpha)$ values) almost consistently
+	improves the prediction rate in (A), but for (B) in some cases causes severe prediction penalties.
+	\item The most obvious example of this is at 1 HL and 256 nodes in (B).
+\end{itemize}
+\medskip
+
+Our "winning" model with 1 HL and 512 nodes preformed rather consistently between training and testing datasets.
+
+\medskip
+\begin{framed}
+\centering
+The Kaggle submission of this model gave a prediction rate of 97.90\%.
+\end{framed}
+\end{frame}
+%-----------------------------------------------------------------
 \begin{frame}{Conclusions}
 
-As expected, our MLP was better suited for handwritten digit recognition.
+Overall, our MLP preformed better on the MNIST dataset
+% As expected, our MLP was better suited for handwritten digit recognition.
+The high-dimensionality of the MNIST dataset provides a difficult hurdle for any RF, however possible improvements could be made to our current model.
+\vspace{3.5mm}
 
-\vspace{3.75mm}
+Data processing to reduce dimensionality and tuning tree depth to add an additional constraint on model complexity could improve the classification accuracy of our RF.
+%Previous studies have shown that models clearly overfit for large depth values.
+Redeveloping our model to consider both the tree depth and the minimum number of leaves would put a stronger constraint on model complexity.
+While more difficult to implement, data processing to reduce dimensionality could be very effective.
+\vspace{3.5mm}
 
-It is possible that data processing to reduce dimensionality would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
-
-\vspace{3.75mm}
-
-One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding additional trees would begin to bridge this gap and further improve it's classification accuracy.
+Even with improvements to our RF, neural network architectures are still better suited for computer vision problems.
+% While it's likely the former changes to our RF will improve it's classification accuracy at a fraction of the MLP's training time, it's classification time will still suffer.
+Overall, computer vision problems are still better left in the hands of NN's and similar models.
 

### 2019-04-23 paper/paper.tex
Paper/Presentation changes

@@
 The MNIST digit database is a classic starting point for individuals wanting to learn the basics of computer vision and get hands on experience with machine learning algorithms.
-The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine [6].
+The dataset is composed of 70,000 28x28 pixel grey-scale images of handwritten numbers between zero and nine [7].
 The pre-flattened Kaggle dataset provides you with 42,000 training examples and 28,000 testing examples.
@@
 This is in part due to it's ease of use, but also due to it's flexibility.
-While in this assignment it is used for classification, it can be also be used in regression problems [2].
-Each tree is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
-Our random forest is a ``highly random'' random forest which splits data with no discretion to promote model volatility.\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
+Random Forest are well apt for both classification and regression problems [2].
+To individuals keen to apply machine learning algorithms but with little computing resources, the speedy training time of RF's is enticing.
+Additionally, for low-dimensional problems overfitting is unlikely with minimal hyper-parameter tuning [2].
 
-These models are generally very fast to train, but depending on the complexity, can suffer from slow run-time performance.
-Generally speaking, overfitting can be negated by simply adding more trees to your model [2].
-This approach eventually results in diminishing returns, and for high-dimensional problems such as digit recognition, is especially impractical [5].
+Each tree of an RF is formed from a bootstrap sample (random sampling with replacement) then grown very similarly to a decision tree.
+Our RF is ``highly random'', meaning it splits data with no discretion to promote model volatility.
+\verb+Sklearn+'s \verb+RandomForestClassifier+ combines the probabilistic prediction of each tree, as this has been shown to preform better than the traditional approach of picking a classification based on a majority vote [4,5].
+
+In the case of high-dimensional problems and real-world applications, the disadvantages of RF's become apparent.
+While RF's training time is consistently minimal in comparison to deep learning architectures, it's run-time performance suffers heavily [6].
+In practical applications, classification speed is often critical to a model's viability.
+In the case of low-dimensional problems, an RF will become more general as more trees are added [2], but this approach eventually results in diminishing returns [5].
+High dimensional datasets frequently have attribute interdependencies, meaning this process of node splitting on a single attribute will be incredibly inefficient [6].
 
@@
 The later can be thought of as a measure of complexity, with lower minimum resulting in higher complexity.
-Once the code was prepared, training took approximately 45 seconds.
+Once the code was prepared, training took approximately 45 seconds per model.
 \begin{figure}[h]
@@
 As expected, our MLP was better suited for handwritten digit recognition.
-It is possible that data processing to reduce dimensionality would help the RF preform better, however computer vision problems are still better left in the hands of NN's and similar models.
-One advantage of the RF is it's training time was approximately $1/60^{th}$ of our MLP's, however adding additional trees would begin to bridge this gap and further improve it's classification accuracy.
+The high-dimensionality of the MNIST dataset provides a difficult hurdle for any RF, however possible improvements could be made to our current model.
+
+Previous studies have shown that models clearly overfit for large depth values.
+Redeveloping our model to consider both the tree depth and the minimum number of leaves would put a stronger constraint on model complexity.
+While more difficult to implement, data processing to reduce dimensionality could be very effective.
+
+While it's likely the former changes to our RF will improve it's classification accuracy at a fraction of the MLP's training time, it's classification time will still suffer.
+Overall, computer vision problems are still better left in the hands of NN's and similar models.
 
@@
 
-[6] Y.
+[6] Xu, B., Huang, J.
+Z., Williams, G., Wang, Q., & Ye, Y.
+(2012).
+\textit{Classifying Very High-Dimensional Data with Random Forests Built from Small Subspaces.} International Journal of Data Warehousing and Mining, 8(2), 44-63.
+doi:10.4018/jdwm.2012040103
+
+[7] Y.
 LeCun \textit{et al} (1995) \textit{Learning Algorithms For Classification: A Comparison On Handwritten Digit Recognition}, in Oh, J.
